\section{Introduction}

A distributed system is any system where nteworked computers (nodes) are physically separated 
and communicates solely by sending messages between them. 

The motivation for a distributed system is the desire to share resources. Examples include 
mobile applications where the user moves around, businesses where resources are located in different locations, 
etc. and ubiquitous computing, in which small cheap devices are embedded in devices and communicate through a network. 

One challenge here is how to make spontaneous connections (service discovery). As the user or the device moves around the network connection does too.

different application poses different chaenges to the distributed system. Real-time application like video and music services must provide a guarantee for the quality and the format. 

A distributed system consist of nodes communicating through messages. A baseline model can be described by 
considering that nodes may leave or enter the network freely. There are three requirements to such a system:

\begin{enumerate}
\item Concurrency: nodes run simultaneously.
\item No global time: the clock in one node is not equal to the clock in another.
\item Independant crash: if one node fails no other nodes should be affected. 
\end{enumerate}

Adding to the challenges of distributed system are 


\begin{enumerate}
\item the heterogenity of devices, of developers skills, .... 	
\item Opennes: Nodes that communicate does so via different interfaces. 
\item Security: hereunder - confidentiality, availability, integrity. Encryption is used to deal with security issues.
\item Scalability: as nodes are added to a system it should scale well i.e., the cost of adding another node should preferably be constant.
\item  
\end{enumerate}

\section{JAXB}

We have used java architecture for XML binding (JAXB) to serialize/deserialize our data objects. 
With JAXB java classes are annotated and the API then transform these objects into valid xml which
can be written to a file.

\section{TCP/UDP}  

TCP 
